Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-02-07 paper/paper.tex
some edits in mnist-like

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. Since $W$ is $\real$- or a proper $\cplx$-Gaussian,
the random variable $
  \tfrac{k \lvert W - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    = \Pr\biggl(
      \frac{k \lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
        \leq \frac{\delta^2 k}{\alpha}
    \biggr)
    = \chi^2_k\biggl(\frac{\delta^2 k}{\alpha}\biggr)
  \,.
\end{equation}
For a generous $\delta = 50\%$ tolerance all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. In our opinion, $\tau = 3$
\citep{molchanov_variational_2017,kingma_variational_2015} retains parameters with randomness
that significantly exceeds the magnitude of the mean, although $\mu$ is typically used as
a stand-in for $W$ when the network runs in inference mode. Given that $q_\theta$ is merely
an approximation, we pick $\tau = -\tfrac12$ to exclude noisy parameters and encourage
higher sparsity.

@@

We conduct a moderately sized numerical experiment on MNIST and similar $28\times 28$
greyscale image datasets to study the compression-performance trade-off of the proposed
$\cplx$-variational dropout on simple architectures.

% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks, and gauge the final performance on the usual test split.
Limiting the size was done primarily to test the regularization strength of sparsification
performed by variational dropout, and to make bulk experimentation faster.

@@

% first experiment: 40-75-40 with fast scheduler
First we try to compare the $\real$-valued networks against their $\cplx$-valued counterparts
with approximately half as many parameters, as suggested by \citet{monning_evaluation_2018}.
In this experiment, ``dense'', ``sparsify'' and ``fine-tune'' stages last for $40$, $75$
and $40$ epochs, respectively. The base learning rate is scaled after $5$, $10$, and $20$
epochs by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$, respectively.

>>>>

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. Since $W$ is $\real$- or a proper $\cplx$-Gaussian,
the random variable $
  \tfrac{k \lvert W - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    % = \Pr\biggl(
    %   \frac{k \lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
    %     \leq \frac{\delta^2 k}{\alpha}
    % \biggr)
    = \chi^2_k\biggl(\frac{\delta^2 k}{\alpha}\biggr)
  \,.
\end{equation}
For a generous $\delta = 50\%$ tolerance all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. In our opinion, $\tau = 3$
\citep{molchanov_variational_2017,kingma_variational_2015} retains parameters with randomness
that significantly exceeds the magnitude of the mean, although $\mu$ is typically used as
a stand-in for $W$ when the network runs in inference mode. Given that $q_\theta$ is merely
an approximation, we pick $\tau = -\tfrac12$ to exclude noisy parameters and encourage
higher sparsity.

@@

We conduct a moderately sized experiment on MNIST and similar $28\times 28$ greyscale
image datasets to study the compression-performance trade-off of the proposed $\cplx$-variational
dropout on simple architectures.
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks, and gauge the final performance on the usual test split.
Limiting the size was done primarily to test the regularization strength of compression,
and to make bulk experimentation faster.

@@

% first experiment: 40-75-40 with fast scheduler
First we compare the $\real$-valued networks against their $\cplx$-valued counterparts with
approximately half as many parameters as suggested by \citet{monning_evaluation_2018}. In
this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base learning
rate is scaled after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.
